Reasoning Graphically about Quantum Computation

### 2008-02-25 papers/reasoning-graphically/paper.tex
hjdfkshfj
git-svn-id: https://isaplanner.svn.sourceforge.net/svnroot/isaplanner/trunk/quantomatic@1372 f8f375fb-4f10-0410-8d47-b6fbdc7ca053

@@

need to define embedding vs strict embedding.

@@

To model the behaviour of quantum systems certain additional equations
must be satisfied.  Each family $(\delta, \epsilon)$ must form a
\emph{classical structure}
\cite{Coecke2006Quantum-Measure,Coecke2006POVMs-and-Naima}:  the pair
$(\delta, \epsilon)$  should form a cocommuative comonoid, the pair
$(\delta^\dag,\epsilon^\dag)$ must form a commutative monoid, and
together they must satisfy the isometry and frobenius equations. In
addition the family $\alpha$ must form an abelian group with $\alpha^\dag
= -\alpha$.  These equations are presented graphically below for the
green family;  the red vertices obey all the same equations
\begin{description}
\item[Comonoid Laws] 
\[
\begin{array}{ccccccccccccccc}
  \inlinegraphic{3em}{comonoid-assoc1}  &=\;&
  \inlinegraphic{3em}{comonoid-assoc2}
  &\qquad\qquad&
  \inlinegraphic{3em}{comonoid-unit1}  &=\;&
  \inlinegraphic{3em}{comonoid-unit2}  &=\;&
  \inlinegraphic{3em}{comonoid-unit3} 
  &\qquad\qquad&
  \inlinegraphic{3em}{comonoid-comm1}  &=\;&
  \inlinegraphic{3em}{comonoid-comm2}
\end{array}
\]
\item[Monoid Laws] 
\[
\begin{array}{ccccccccccccccc}
  \inlinegraphic{3em}{monoid-assoc1}  &=\;&
  \inlinegraphic{3em}{monoid-assoc2}
  &\qquad\qquad&
  \inlinegraphic{3em}{monoid-unit1}  &=\;&
  \inlinegraphic{3em}{monoid-unit2}  &=\;&
  \inlinegraphic{3em}{monoid-unit3} 
  &\qquad\qquad&
  \inlinegraphic{3em}{monoid-comm1}  &=\;&
  \inlinegraphic{3em}{monoid-comm2}
\end{array}
\]
\item[Isometry]
\[
\begin{array}{ccc}
  \inlinegraphic{3.5em}{isometry1}  &=\;&
  \inlinegraphic{3.5em}{isometry2}  
\end{array}
\]
\item[Frobenius]
  \[
  \begin{array}{ccccc}
    \inlinegraphic{3em}{frobenius1}  &=\;&
    \inlinegraphic{3em}{frobenius2} &=\;&
    \inlinegraphic{3em}{frobenius3}
  \end{array}
  \]
  \item[Abelian Unitary Group]
    \[
    \begin{array}{ccccccccccccccc}
      \inlinegraphic{2.2em}{group1}  &:=\;&
      \inlinegraphic{2.2em}{group2}  &=\;&
      \inlinegraphic{2.2em}{group3} 
      &\qquad\qquad&
      \left( \inlinegraphic{2.2em}{group4}\!\right)^\dag  &=\;&
      \inlinegraphic{2.2em}{group5} 
      &\qquad\qquad&
      \inlinegraphic{3.5em}{group6}  &=\;&
      \inlinegraphic{3.5em}{group7}  &=\;&
      \inlinegraphic{3.5em}{group8} 
    \end{array}
    \]
\end{description}

@@

>>>>

\begin{definition}
  A \emph{directed graph} consists of a 4-tuple $(N,E,s,t)$ where $V$
  and $E$ are sets, respectively of \emph{nodes} and \emph{edges}, 
\end{definition}

@@

need to define embedding and subgraph

@@

To model the behaviour of quantum systems certain additional equations
must be satisfied.  Each family $(\delta, \epsilon)$ must form a
\emph{classical structure}
\cite{Coecke2006Quantum-Measure,Coecke2006POVMs-and-Naima}:  the pair
$(\delta, \epsilon)$  should form a cocommuative comonoid, the pair
$(\delta^\dag,\epsilon^\dag)$ must form a commutative monoid, and
together they must satisfy the isometry and frobenius equations. In
addition the family $\alpha$ must form an abelian group with $\alpha^\dag
= -\alpha$.  These equations are presented graphically below for the
green family;  the red vertices obey all the same equations
\begin{description}
\item[Comonoid Laws] 
\[
\begin{array}{ccccccccccccccc}
  \inlinegraphic{3em}{comonoid-assoc1}  &=\;&
  \inlinegraphic{3em}{comonoid-assoc2}
  &\qquad\qquad&
  \inlinegraphic{3em}{comonoid-unit1}  &=\;&
  \inlinegraphic{3em}{comonoid-unit2}  &=\;&
  \inlinegraphic{3em}{comonoid-unit3} 
  &\qquad\qquad&
  \inlinegraphic{3em}{comonoid-comm1}  &=\;&
  \inlinegraphic{3em}{comonoid-comm2}
\end{array}
\]
\item[Monoid Laws] 
\[
\begin{array}{ccccccccccccccc}
  \inlinegraphic{3em}{monoid-assoc1}  &=\;&
  \inlinegraphic{3em}{monoid-assoc2}
  &\qquad\qquad&
  \inlinegraphic{3em}{monoid-unit1}  &=\;&
  \inlinegraphic{3em}{monoid-unit2}  &=\;&
  \inlinegraphic{3em}{monoid-unit3} 
  &\qquad\qquad&
  \inlinegraphic{3em}{monoid-comm1}  &=\;&
  \inlinegraphic{3em}{monoid-comm2}
\end{array}
\]
\item[Isometry]
\[
\begin{array}{ccc}
  \inlinegraphic{3.5em}{isometry1}  &=\;&
  \inlinegraphic{3.5em}{isometry2}  
\end{array}
\]
\item[Frobenius]
  \[
  \begin{array}{ccccc}
    \inlinegraphic{3em}{frobenius1}  &=\;&
    \inlinegraphic{3em}{frobenius2} &=\;&
    \inlinegraphic{3em}{frobenius3}
  \end{array}
  \]
\item[Compact Structure] 
  \[
  \begin{array}{ccccccc}
    \inlinegraphic{2.5em}{compact1}  &=\;&
    \inlinegraphic{1.3em}{compact2}  
    &\qquad\qquad&
    \inlinegraphic{2.5em}{compact3}  &=\;&
    \inlinegraphic{1.3em}{compact4}  
  \end{array}
  \]
  \item[Abelian Unitary Group]
    \[
    \begin{array}{ccccccccccccccc}
      \inlinegraphic{2.2em}{group1}  &:=\;&
      \inlinegraphic{2.2em}{group2}  &=\;&
      \inlinegraphic{2.2em}{group3} 
      &\qquad\qquad&
      \left( \inlinegraphic{2.2em}{group4}\!\right)^\dag  &=\;&
      \inlinegraphic{2.2em}{group5} 
      &\qquad\qquad&
      \inlinegraphic{3.5em}{group6}  &=\;&
      \inlinegraphic{3.5em}{group7}  &=\;&
      \inlinegraphic{3.5em}{group8} 
    \end{array}
    \]
  \item[Bilinearity] 
    \[
    \begin{array}{cccccccccccc}
      \inlinegraphic{3.5em}{alpha-commute1}  &=\;&
      \inlinegraphic{3.5em}{alpha-commute2}  &=\;&
      \inlinegraphic{3.5em}{alpha-commute3} 
      &\qquad\qquad&
      \inlinegraphic{3.5em}{alpha-commute4}  &=\;&
      \inlinegraphic{3.5em}{alpha-commute5}  &=\;&
      \inlinegraphic{3.5em}{alpha-commute6} 
    \end{array}
    \]
\end{description}
Additionally, the two families satisfy equations which make them a
\emph{scaled bialgebra}.
\begin{description}
\item[Bialgebra Laws] Let  $\inlinegraphic{1em}{bialgebra1} := 
      \inlinegraphic{1.5em}{bialgebra2}$;  then:
  \[
  \begin{array}{ccccccccccccccc}
      \inlinegraphic{3.5em}{bialgebra3}  &=\;&
      \inlinegraphic{3em}{bialgebra4}
      \qquad\qquad
      \inlinegraphic{2.5em}{bialgebra5}  &=\;&
      \inlinegraphic{1.5em}{bialgebra6}
      \qquad\qquad
      \inlinegraphic{2.5em}{bialgebra7}  &=\;&
      \inlinegraphic{1.5em}{bialgebra8} \\ \\
      \inlinegraphic{3.5em}{bialgebra9}  &=\;&
      \inlinegraphic{3em}{bialgebra10}
      \qquad\qquad
      \inlinegraphic{2.5em}{bialgebra11}  &=\;&
      \inlinegraphic{1.5em}{bialgebra12}
      \qquad\qquad
      \inlinegraphic{2.5em}{bialgebra13}  &=\;&
      \inlinegraphic{1.5em}{bialgebra14} 
  \end{array}
  \]
\end{description}

@@

Interaction nets and interaction combinators (Lafont and others)
